---
title: 'Statistical Analysis of Climate Impact on Global Agriculture Data'
author: Allison Herr and Kelvin Lartey 
date: \today
output: 
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
geometry: margin=0.85in
fontsize: 12pt
header-includes:
- \usepackage{setspace}\onehalfspacing
editor_options: 
  markdown: 
    wrap: 72
---

\listoffigures
\listoftables

\newpage

```{r, message = FALSE, warning = FALSE}
# set your global options here and load your packages
knitr::opts_chunk$set(fig.width = 10, fig.height = 5, echo = TRUE, eval = TRUE)
library(knitr)
library(tidyverse)
library(readr)
climate_change_impact_on_agriculture_2024 <- read_csv("climate_change_impact_on_agriculture_2024.csv")
head(climate_change_impact_on_agriculture_2024)
```


# Introduction

This project was done to fulfill the requirements of MAT317:
Introduction to Statistical Learning. Allison Herr and Kelvin Lartey.              
performed x% and 100 - x% of the work, respectively.

Describe here what are you trying to analyze and why would anyone care
about your work. Two paragraphs are enough.

Our data set is looking at the 








# Climate Zones
Creating climate zones based on temperature and precipitation
```{r}


Climate_newdata <- climate_change_impact_on_agriculture_2024 %>%
  mutate(
    climate_zone = case_when(
      Average_Temperature_C <= 0 ~ "Polar_zone",
      Total_Precipitation_mm <= 300 ~ "Dry_zone",
      Average_Temperature_C >= 20 ~ "Tropical_zone",
      Average_Temperature_C >= 10 & Average_Temperature_C < 20 ~ "Temperate_zone",
      Average_Temperature_C > 0 & Average_Temperature_C < 10 ~ "Continental_zone"
    )
  )

head(Climate_newdata)

```


# Average crop yield by climate zone

```{r}
ggplot(Climate_newdata, aes(x = climate_zone, y = Crop_Yield_MT_per_HA)) +
  geom_boxplot() +
  labs(
    title = "Crop Yield by Climate Zone",
    x = "Climate Zone",
    y = "Crop Yield (MT per HA)"
  ) +
  theme_minimal()
```
# Crop yield trend over time; over time is crops improving or not
Hello, check for commit
```{r}
yield_by_year <- Climate_newdata %>%
  group_by(Year) %>%
  summarise(
    mean_yield = mean(Crop_Yield_MT_per_HA, na.rm = TRUE),
    n = n()
  )

yield_by_year

ggplot(yield_by_year, aes(x = Year, y = mean_yield)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Crop Yield Over Time",
    x = "Year",
    y = "Crop Yield (MT per HA)"
  ) +
  theme_minimal()
```


#Crop yield by the crop type


```{r}
ggplot(Climate_newdata, aes(x = Crop_Type, y = Crop_Yield_MT_per_HA)) +
  geom_boxplot() +
  labs(
    title = "Crop Yield by Crop Type",
    x = "Crop Type",
    y = "Crop Yield (MT per HA)"
  ) +
  theme_minimal()
```


#Extreme weather events and crop yield
Extreme weather effect, barely changes the mean yield. That is intresting
```{r}
Climate_newdata %>%
  group_by(Extreme_Weather_Events) %>%
  summarise(
    mean_yield = mean(Crop_Yield_MT_per_HA, na.rm = TRUE),
    n = n()
  )
```


#Irrigation access and crop yield
There is no correlation between irrigation and crop yield.
```{r}

Climate_newdata<- Climate_newdata%>%
  rename(Irrigation_Access_per = 'Irrigation_Access_%')


#change the Irrigation_Access_% to Irrigation_Access_per, but there was something weird happeing so if anything that might be an issue
ggplot(Climate_newdata, aes(x = Irrigation_Access_per, y = Crop_Yield_MT_per_HA)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Crop Yield vs Irrigation Access",
    x = "Irrigation Access (%)",
    y = "Crop Yield (MT per HA)"
  ) +
  theme_minimal()


cor(
  Climate_newdata$`Irrigation_Access_per`,
  Climate_newdata$Crop_Yield_MT_per_HA,
)
```


# Comparing crop yield across different climate zone and crop type all together

```{r}
ggplot(Climate_newdata, aes(x = climate_zone, y = Crop_Yield_MT_per_HA, fill = Crop_Type)) +
  geom_boxplot() +
  labs(
    title = "Crop Yield by Climate Zone and Crop Type",
    x = "Climate Zone",
    y = "Crop Yield (MT per HA)"
  ) +
  theme_minimal()
```


```{r}
Climate_newdata %>%
  group_by(climate_zone, Crop_Type) %>%
  summarise(
    mean_yield = mean(Crop_Yield_MT_per_HA, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  arrange(climate_zone, desc(mean_yield))
```



# Overall Correltion data

```{r}
numeric_data <- Climate_newdata %>%
  select(
    Year,
    Average_Temperature_C,
    Total_Precipitation_mm,
    CO2_Emissions_MT,
    Crop_Yield_MT_per_HA,
    Extreme_Weather_Events,
    `Irrigation_Access_per`
  )

cor(numeric_data)
```



## Research questions

State here your specific regression and classification research
questions. Use specific variable names that you will use in your
analyses.

## Data set description

Our data set, Global Agriculture Climate Change impact was found on
Kaggle
(<https://www.kaggle.com/datasets/talhachoudary/global-agriculture-climate-impact-dataset>)
What do you know about the data? How was it collected? When was it
collected? Who collected it? This data set contains 15 variables, the
some of the most relevant variables are Extreme weather, a categorical
variable that depicts different weather events What are the relevant
variables?
