---
title: "Comments for your project"
author: Prof. Draguljić
date: \today
fontsize: 12pt
geometry: margin=0.7in
output: pdf_document
header-includes:
- \usepackage{setspace}\onehalfspacing
---

\tableofcontents
\listoffigures
\listoftables

\newpage



See the RMarkdown file for how to easily create Table of contents, 
list of figures, and list of tables. 

What follows is a list of common mistakes that I compiled from different 
courses. Please address the mistakes in your work to make it more
professional.

Stop by my office if ask a question beofre you turn in substandard work. 

# Setting global R options

To set global options like figure size, printing of R code, etc. use this
R command. 

```{r, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 10, fig.height = 5, echo = TRUE, eval = TRUE)
```

# Suspending warning messages

If you load the tidyverse package you get all sorts of warnings. They are 
not necesarily bad. It's just R telling what it's doing. 

```{r}
library(tidyverse)
```

These comments should not be in your final report. To supress them
use *message = FALSE, warning = FALSE* options in the R chunk where you 
load the package. 

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
```

# General comments

## Read your report

Read your final report before you submit it! And read it in your pdf
file and not your rmd file. I will be reading your pdf file, so if 
something looks good in rmd file but not in pdf file, then it 
does not look good. 

Read. Your. Final. Submission!

More than once. 

## Provide detailed answers

For any question (where applicable) provide written and 
visual answers (figures, tables, etc.) The written and visual 
answers should be stand-alone. That is, after reading your written
response I should be able to imagine what that scatterplot or
boxplot looks like. 

## Creating figures

Figure \ref{fig:q1a} shows a bad looking plot. The same plot that looks 
decent is shown in Figure \ref{fig:q1b}. Note how I referenced the 
figures here using LaTeX. Not how I also added figure labels that also 
show up in the list of figures. 

```{r,  fig.cap="\\label{fig:q1a} The scatterplot of Sepal.Length and Petla.lenghts"}
plot(iris$Sepal.Length, iris$Petal.Length, cex.lab = 0.75, cex.axis = 0.75, 
     main = "Petal vs. Sepal length")
```



```{r,  fig.cap="\\label{fig:q1b} The relationship between sepal and petal lenghts", fig.width = 6, fig.height = 5}
plot(iris$Sepal.Length, iris$Petal.Length, 
     xlim = c(4, 8), las = 1, ylim = c(1, 7),
     xlab = "Sepal length (cm)", ylab = "Petal length (cm)",
     cex.lab = 1.1, cex.axis = 1.1)
```

Here are some comments on the bad plot. 

\begin{itemize}
   \item Axis and tickmark labels are too small. 
   \item Both axis ranges are too small, i.e. they do not cover all the 
      data. 
   \item Since both variables are on the same scale, the plot should be more
      square than rectangular. 
   \item Don't put a main title in your plot. That's what the figure label is for. 
   \item Figure label 
   \begin{itemize}
      \item There are typos in there.
      \item Never user R's variable names in your labels or, for that matter,
         in the text. Never! These are random names and might have no meaning
         relevant at all. 
      \item I see it's a scatterplot. No need to tell me that. 
      \item Don't capitalize ``random'' words. Start with a capital letter
         and capitalize what you would capitalize otherwise. 
   \end{itemize}
   \item Tickmark labels aren't rotated properly on the y-axis. 
   \item Units are missing. 
\end{itemize}

Keep your figures, generally, in the same format, i.e. the same label size, 
same tickmark size, etc., in the whole document.  

Don't use colors that cannot be clearly seen in your plots. In particular, I 
see a lot of yellow plot symbols. They just tended to blur together 
on the page. So, yellow makes a particularly bad choice. 

You don't need the legend if there is only one symbol or 
line on the plot. Two or more, add a legend.

## Creating tables

- Add table labels. Don't make your labels too technical.  

```{r}
summaryStatIris <- iris %>% 
  group_by(Species) %>%
  summarise(n = n(), 
            avg = round(mean(Sepal.Length), 2), 
            sd = round(sd(Sepal.Length), 2), 
            lcb = round(avg - qt(0.975, n - 1)*sd/sqrt(n), 2), 
            ucb = round(avg + qt(0.975, n - 1)*sd/sqrt(n), 2))
```


This is a bad table: 

```{r}
summaryStatIris
```

Table \ref{tbl:sumStats} shows the summary statistics of the sepal length for 
each species. This table looks good. It is properly formatted and it has 
a table label that shows up in list of tables at the beginning. 

```{r}
library(knitr)
kable(summaryStatIris, 
      col.names = c("Species", "n", "Average", "Std. deviation", 
                    "95% CI lower bound", "95% CI upper bound"), 
      caption = "\\label{tbl:sumStats} The summary statistics for the sepal length by species")
```

## Hard-coding the values 

Do not hard-code values in your code. Eventually, this will be a bad news!

For example

```{r}
mean(iris$Petal.Length)
mean(iris$Sepal.Length)
if(3.758 < 5.843333){
  junkVar <- iris$Petal.Length - (5.843333 - 3.758)
}
mean(junkVar)
```

should be coded as 

```{r}
avgP <- mean(iris$Petal.Length)
avgS <- mean(iris$Sepal.Length)
if(avgP < avgS){
  junkVar <- iris$Petal.Length - (avgS - avgP)
}
mean(junkVar)
```

With the proper coding, if your data set changes, your code will still work. If you 
hard-code in the values, then you will have to go around and fix all those. 
You will miss some for sure. 

## Wrapping your R code

Keep your code on the page and in the shaded box. 

```{r,  eval = FALSE}
plot(iris$Sepal.Length, iris$Petal.Length, xlim = c(4, 8), las = 1, ylim = c(1, 7), xlab = "Sepal length (cm)", ylab = "Petal length (cm)", cex.lab = 1.1, cex.axis = 1.1)
```

Nobody can read the code above. See Figure \ref{fig:q1b} code for proper and
readable formatting. 

## Writing up the results 

When you are describing a distribution, describe this distribution in terms of 
center or typical value (mean, median, or some other appropriate statistic), 
spread (usually standard deviation, IQR, or range), outliers, skewness, modality, 
etc. 

When you're asked to compare two distributions, most often it makes sense to 
compare the average values. Saying something like 

"on average, our data indicates that
early class is getting higher scores than the afternoon class" 

is much better than saying

"there is some overlap between boxplots for group A and group B."



## Write in context of the problem 

All your written explanations, interpretation, etc. should be in the 
context of a problem! They should also try to be as non-technical as 
possible. 

While saying something like 

"the distribution of the scores is skewed right"

is correct, saying something like

"majority of the countries have low GBGR scores indicating low
acceptance of LGBTQ community"

will be much more appreciated by your non-statistical colleague. 


## Using R in the text 

You can use R within your text. For example, suppose you want to say 
what the average sepal length in your data set is. You could do this 
by calculating the height first via

```{r}
mean(iris$Sepal.Length)
```

and then, after copying and pasting the average height, saying 
something along the line of 

"The average child's height is 5.84 cm."

Another option is to directly calculate this height in your text as in 

"The average child's height is `r round(mean(iris$Sepal.Length), 2)` inches."

The above line reads 

The average child's height is **r mean(iris$Sepal.Length), 2)** inches.

where the bolded *r mean(iris$Sepal.Length), 2)* is placed into ` ` so that Markdown
knows that it is an R command and that it should execute it. I bolded it here
for the illustration purposes. 

<!--
## 11. 

Consider the two plots below. 

```{r, eval = FALSE, echo = FALSE}
set.seed(200)
x <- runif(100, 0, 1)*100

y1 <- x + rnorm(100, 0, 20)
y2 <- x + rnorm(100, 0, 4)

par(mfrow = c(1, 2))
plot(x, y1)
abline(a = 0, b = 1)
plot(x, y2)
abline(a = 0, b = 1)
```

Both plots show a pretty strong relationship between x and y. As a matter of fact, 
both sets of data points were generated from $y = x$ relationship. The difference 
is that, in the plot on the left, I added more noise about the line than in 
the plot on the right. While both plots show pretty strong relationship between the
variables, the variability is much larger in the plot on the left. 
-->

## Additional comments

- Don't use the R variable names in your written part. If you make a 
histogram of via *hist(myData$hgb)*, you should not use $hgb$ in the 
text. If, for example, *hgb* represents the hemoglobin blood levels, 
then you should use hemoglobin blood levels in your document. 

- To interpret coefficients say something like 

When the stove is turned off, our model predicts that 450.19 seconds are 
needed for 1200 g of snow to melt on average.

- Write down you estimated model:

The straight line relationship between the stove setting and the time
needed to melt 120 g of snow is given by 

\[temperature = 450.19 - 234\times stove~setting.\]

<!--
- var(response) cannot be reduced by a model. 
    + "Model doesn't account for variability in response. It hits only a few values."
    isn't a clear statement. 
-->

- Run the spell check! Each misspelled word is penalized. The spell check 
button is next to the knit button and it looks like ABC with a check below them. 

- Round your values to something reasonable.
- Show your calculations, i.e. type them out. For example

\begin{eqnarray*}
\Delta y &=& y(x+1) - y(x) \\
   &=& \beta_0 + \beta_1(x+1) + \beta_2(x+1)^2 - \beta_0 + \beta_1x + \beta_2x^2 \\
   & & \vdots \\
   &=& \beta_1 + \beta_2 + 2\beta_2x  
\end{eqnarray*}

- Write out your derivative in LaTeX. For example

\[\frac{\mathrm{d}}{\mathrm{d} x} y = -65e^{-x}\]

- To add space between two words in LaTeX's math environment use ~ as in 

\[stove setting\] vs \[stove~setting\]


 
- Don't start your answer with "answer:" but do start them on a new line. 
- Use LaTeX where you can. 
