---
title: 'Statistical Analysis of ### Data'
author: Ben Franklin and John Marshall
date: \today
output: 
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
geometry: margin=0.85in
fontsize: 12pt
header-includes:
- \usepackage{setspace}\onehalfspacing
---

\listoffigures
\listoftables

\newpage

```{r, message = FALSE, warning = FALSE}
# set your global options here and load your packages
knitr::opts_chunk$set(fig.width = 10, fig.height = 5, echo = TRUE, eval = TRUE)
library(knitr)
library(tidyverse)
```

# Introduction

This project was done to fulfill the requirements of MAT317: Introduction
to Statistical Learning. Ben Franklin and John Marshall performed x% and 
100 - x% of the work, respectively. 

Describe here what are you trying to analyze and why would anyone care
about your work. Two paragraphs are enough. 

## Reserach questions

State here your specific regression and classification 
research questions. Use specific variable names that you will 
use in your analyses. 

## Data set desription

Where did you get your data set? Give a website if you have one. 
What do you know about the data? How was it collected? When was it 
collected? Who collected it? What are the relevant variables?


# Statistical Methods

## Regression 

What are the two statistical techniques you plan to use to answer your 
regression question? Give details here. Name variables, 
write out models (using $\beta_i$'s), let me know if you're using CV,
backward selection, etc. Write formulas. Explain why you chose
to use these two techniques. 

Make a histogram of your Y variable and describe it. Make a 
scatterplot of Y and an X that might be important and describe
that relationship. Give me a feel for your data.  

Make a summary stats table of your response variable (n, average, min, 
Q1, Q2, Q3, and max). See Table \ref{tbl:sumStats} for an example. 

## Classification

What are the two statistical techniques you plan to use to answer your 
regression question? Give details here. Name variables, 
write out models (using $\beta_i$'s), let me know if you're using CV,
backward selection, etc. Write formulas. Explain why you chose
to use these two techniques. 

Make a comparative boxplot of your Y variable and describe it. Make a 
boxplot / scatterplot of Y by X that might be important and describe
that relationship. Give me a feel for your data.  

Make a summary stats table of your response variable. How many observations
in each category?

# Results

## Regression 

Write estimated final models. Give details. Interpret 
models and/or coefficients. What do your models say? Do the two 
models send the same message? What are the important inputs? 

Make plots of your models if possible. 


## Classification

Write estimated final models. Give details. Interpret 
models and/or coefficients. What do your models say? Do the two 
models send the same message? What are the important inputs? 

Make plots of your models if possible. 


```{r,  fig.cap="\\label{fig:q1b} The relationship between sepal and petal lenghts", fig.width = 6, fig.height = 5}
plot(iris$Sepal.Length, iris$Petal.Length, 
     xlim = c(4, 8), las = 1, ylim = c(1, 7),
     xlab = "Sepal length (cm)", ylab = "Petal length (cm)",
     cex.lab = 1.1, cex.axis = 1.1)
```


```{r}
summaryStatIris <- iris %>% 
  group_by(Species) %>%
  summarise(n = n(), 
            avg = round(mean(Sepal.Length), 2), 
            sd = round(sd(Sepal.Length), 2), 
            lcb = round(avg - qt(0.975, n - 1)*sd/sqrt(n), 2), 
            ucb = round(avg + qt(0.975, n - 1)*sd/sqrt(n), 2))


kable(summaryStatIris, 
      col.names = c("Species", "n", "Average", "Std. deviation", 
                    "95% CI lower bound", "95% CI upper bound"), 
      caption = "\\label{tbl:sumStats} The summary statistics for the sepal length by species")
```

# Conclusions

What did you learn? What else would you have wanted to know but couldn't?
Is something looking weird / surprising / unexpected?

Don't use formulas here or be too statistical. This section is for the 
wider audience. 

# Appendix A

If you need it. 